Non-word repetition in Yélî Dnye

### 2020-12-01 manuscript.Rmd
continued methods edits

A Praat script was written to randomize this list 20 times, and split it into two sublists, to generate 40 different elicitation sets. The 40 elicitation sets are available online from https://osf.io/5qspb. The split had the following constraints:

@@

We tried to balance three desiderata: That children would not be unduly exposed to the items before they themselves had to repeat them; that children would feel comfortable doing this task with us; and that the community would feel safe with us doing this task with their children. Moreover, there were also some logistic constraints in terms of the space availability. As a result, the places where elicitation happened varied across the hamlets.

We visited four different hamlets once, and attempted to test all eligible children present at the time, to prevent the items 'spreading' through hearsay. In the first hamlet, we tested children in five different places, with some children being tested inside a house and others tested on the veranda. The complete list of places and the ways in which they met the desiderata mentioned above can be found in the raw data, available from online supplementary materials.

The child was donned a headset (a Shure SM10A or WH20 XLR with a dynamic microphone on a headband, most children using the former), recorded into the left channel into a Tascam DR40x digital recorder. For most children, the headset could not stay comfortably on the child's head, and thus it was placed on the child's shoulders, with the microphone carefully placed close to the child's mouth. A local informant sat next to the child, to would provide the instructions and, if needed, coach the child to make sure, using the three practice items as well as real words, that they understood that the task was to repeat the items precisely without changing anything. An experimenter (the first author) delivered the elicitation stimuli to the local informant and the child over headphones.

The first phase was making sure the child understood the task. This was explained orally and the first training item was presented. Often, children froze and did not say anything. If this happened, then we followed this procedure. First, the informant insisted. If the child still did not say anything, the informant asked the child to repeat a real word, and another, and another. If the child could repeat these correctly, then we provided the recorded training item over headphones again. Most children successfully started repeating the items presented over headphones at this point; a few further needed the local informant to model the behavior (i.e., they would hear the item again, and she would say it; then we would play it again, and ask the child to say it). A small minority still failed to repeat the item after hearing it over headphones. If that occurred, we tried with the second training item, at which point some children got it and could continue. A small minority, however, failed to repeat this one, as well as the third training item, in which case we stopped the test altogether (see Participants section for exclusions).

NWR studies vary in whether children are provided with several opportunities to hear and say the item. To have a fixed and clear procedure, we decided that items other than the inital three training ones would not be repeated unless the child made an attempt to produce them. If this attempt was judged correct by the local informant, then the experimenter would move on to the next item (whispering this over a separate headset that was recording onto the right track of the same Tascam). If the local informant heard a deviation, she indicated to the experimenter that the item needed to be repeated, and up to 5 attempts were allowed.

Whenever siblings from the same family were tested, an attempt was made to test first the older and then the younger child, and always on different elicitaiton sets.

@@

A script was written to randomize all tokens from all children, pairing each with the auditory target the child had been provided with.  A native research assistant then listened through all productions of a given target (randomized across children and repetition order), and made a judgment of whether the item was correctly or incorrectly repeated. She additionally orthographically transcribed exactly what the child said, providing some examples of the types of errors children in general make (without making specific reference to Yélî sounds or the items in the elicitation sets).

@@

Some NWR studies employ phoneme-based scoring in addition to or instead of word-level scoring [e.g., @cristia2020infant]. We scored items in terms of the number of phonemes that could be aligned across the target and attempt, divided by the number of phonemes of whichever item was longer (the target or the attempt). Although previous work does not use distance metrics, we additionally report those. For instance, recall our example of an English target being "bilik", and imagine a response "bilig". This would receive the following scores: at the whole item level this would receive a score of zero (because the repetition is not exact); at the phoneme level it would receive a score of 80% (4 out of 5 phonemes repeated exactly); and its normalized phone-based Levenshtein distance is 20% (20% of phonemes were substituted or deleted).  
<!--M2A: A concrete scoring example would be a good idea here-->
<!-- a2m added! actually, I haven't changed anything but I've been hesitant about using levenshtein's distance because it's the complement of phoneme based scoring -- so shall we drop it?  -->
<!-- M2A: three thoughts on this, which you can take as you like:
1) our audience might make this determination. segment-match score (phoneme-level accuracy) is probably more straightforward and might therefore be preferred by psychies
2) LD has the advantage of making D/I/S relevant as a next step in the analysis
3) we should definitely choose one, I agree, and report the other in supp mat tables
-->

Additionally, we looked up each phoneme in each target word in the Phoible database [@phoible]. We entered the number of languages in which each phoneme was found, as well as what percentage of languages in the database that represented. We could not find a small number of phonemes in the database, and treat those as NA values that do not contribute to the target-level global average phone cross-linguistic frequency, although an alternative would be to attribute them a very small number.

Finally, for describing children's patterns of errors, all repetitions of a given target were taken into account. We describe the proportion of items where the change resulted in a real word (semantic errors); as well as deletions, additions, and substitutions.
<!-- a2m in the results, I put analyses based on all as well as just on first attempt. If we do just the latter, then come back to change this para -->

>>>>

A Praat script [@Praat] was written to randomize this list 20 times, and split it into two sublists, to generate 40 different elicitation sets. The 40 elicitation sets are available online from https://osf.io/5qspb. The split had the following constraints:

@@

In adapting the typical NWR procedure for this context, we balanced three desiderata: That children would not be unduly exposed to the items before they themselves had to repeat them (i.e., from other children who had participated); that children would feel comfortable doing this task with us; and that community members would feel comfortable having their children do this task with us.

We ran children in batches, testing within a handful of hamlet clusters spread across the northeastern region of the island. Because space availability was limited in different ways from hamlet to hamlet, the places where elicitation happened varied across testing sites. We tested in four different sites, only making a single visit to each, conducting back-to-back testing of all eligible children present at the time of our visit in order to prevent the items 'spreading' between children through hearsay. In the first hamlet, we tested children in five different places, with some children being tested inside a house and others tested on the veranda. The complete list of places and the ways in which they met the desiderata mentioned above can be found in the raw data, available from online supplementary materials. Whenever children living in the same household were tested, we tried to test children in age order, from oldest to youngest, to minimize intimidation for younger household members, and always using different elicitaiton sets.

We fitted the child with a headset microphone (Shure SM10A or WH20 XLR with a dynamic microphone on a headband, most children using the former) that fed into the left channel of a Tascam DR40x digital audio recorder. The headsets were designed for adult use and could not be comfortably seated on many children's heads without a more involved adjustment period. To minimize adjustment time, which was uncomfortable for some children given the proximity of the experimenter and equipment, we placed the headband on children's shoulders in these cases, carefully adjusting the microphone's placement so that it was still close to the child's mouth. A research assistant who spoke Yélî Dnye natively sat next to the child throughout the task to provide instructions and, if needed, encouragement. The research assistant coached the child throughout to task to make sure that they understood what they were expected to do. An experimenter (the first author) delivered the pre-recorded stimuli to the research assistant and the child over headphones.

The first phase of the experiment involved making sure the child understood the task. We explained the task orally presented the first practice item. At this point, many children did not say anything in response, at which point we initiated the following procedure: First, the assistant insisted the child make a response. If the child still did not say anything, the assistant said a real word and then asked the child to repeat it, then another and another. If the child could repeat real words correctly, we provided the first training item over headphones again for children to repeat. Most children successfully started repeating the items at this point, but a few needed further help. In this case, the assistant modeled the behavior (i.e., the child and assistant would hear the item again, and the assistant would repeat it; then we would play the item again and ask the child to repeate it). A small minority of children still failed to repeat the item at this point. If so, we tried again with the second training item, at which point some children got it and could continue. A fraction of the remaining children, however, failed to repeat this second training item, as well as the third one, in which case we stopped testing altogether (see Participants section for exclusions).

The second phase of the experiment involved running the child through the list of test items randomly assigned to them. This was done in the same manner as the practice items: the stimulus was played over the headphones, and then the child repeated it aloud. NWR studies vary in whether children are allowed to hear and/or repeat the item more than one time. We had a fixed procedure for the test items (i.e., the non-practice items) in which the child was alowed to make a further attempts if their first attempt was judged erroneous in some way by the assistant. The procedure worked as follows: When the child made an attempted a repeat, the assistant indicated to the experimenter whether the child's production was correct or not. If correct, the experimenter would whisper this note of correct repetition into a separate headset that fed into the right channel of the same Tascam and the group moved on to the next item. If not, the child was allowed to try again, with up to five attempts allowed before moving on to the next item. Children were not asked to make repetitions if they did not produce a first attempt.

In total, test sessions took approximately XX minutes on average, with XX of those minutes attributed to practice and XX to the actual test list.
<!--M2A: Alex, could you please estimate the above XX cases and fill in?-->

@@

The first author then annotated the onset and offset of all children's productions from the audio recording using Praat audio annotation software [@Praat],
<!--M2A: Alex, I added a bib entry for Praat but it's the most recent version. Could you please correct it to the version (approximately if not precisely) that you used?-->
then wrote and ran a script to extract these tokens, pairing them with their original auditory target stimulus, and writing these audio pairs out to short .wav files. The assistant then listened through all these productions, one target item at a time, with productions for each target item presented in a random order across children and repetitions. The assistant indicated in a notebook whether each production was a correct or incorrect repetition and orthographically transcribed the production, noting when the child uttered a recognizable word or phrase and adding the translation equivalent of that word/phrase into English. The assistant also provided some general examples of the types of errors children made without making specific reference to Yélî sounds or the items in the elicitation sets.

@@

In addition to a binary word-level exact repetition score we scored items in terms of the number of phonemes that could be aligned across the target and attempt, divided by the number of phonemes of whichever item was longer [the target or the attempt; see also @cristia2020infant]. Although previous work does not use distance metrics, we additionally report these. For instance, recall our example of an English target being /bilik/ with an imagined response [bilig]. We would score this response as follows: at the whole item level this production would receive a score of zero (because the repetition is not exact); at the phoneme level this production would receive a score of 80% (4 out of 5 phonemes repeated exactly); and the phone-based Levenshtein distance for this production is 20% (because 20% of phonemes were substituted or deleted).  
<!--M2A: A concrete scoring example would be a good idea here-->
<!-- a2m added! actually, I haven't changed anything but I've been hesitant about using levenshtein's distance because it's the complement of phoneme based scoring -- so shall we drop it?  -->
<!-- M2A: three thoughts on this, which you can take as you like:
1) our audience might make this determination. segment-match score (phoneme-level accuracy) is probably more straightforward and might therefore be preferred by psychies
2) LD has the advantage of making D/I/S relevant as a next step in the analysis
3) we should definitely choose one, I agree, and report the other in supp mat tables-->

<!--M2A: So what's the final determination here, keeping in mind we're now aiming for LDR?-->

Additionally, we estimated the typological frequency of all phonological segments used in the target items using the PHOIBLE cross-linguistic phonological inventory database [@phoible]. For each phone in our task, we extracted the number and percentage of languages noted to have that phone in its inventory. While PHOIBLE is an unprecedentedly comprehensive database, with phonological inventory data for over 2000 languages at the time these data were analyzed, it is of course still far from complete, so estimated frequencies should be taken with a pinch of salt. In both PHOIBLE and UPSID---another large phonemic inventory database---note that nearly half of the segment types only appear in a single language (Steven Moran, personal communication; see also [web.phonetik.uni-frankfurt.de/upsid_info.html](http://web.phonetik.uni-frankfurt.de/upsid_info.html)). It is therefore unsurprising that we did not find three of our segments (/lβʲ/, /ṭp/, and /tp/) in the database. Following the general pattern that many phones only appear once cross-linguistically, even in large databases, we treat our "missing" phones as having a frequency of 1, with a (rounded) percentage of 0%.

Finally, when describing children's patterns of errors, we take all repetitions of a given target into account. We describe the proportion of items where the change resulted in a real word (semantic errors); as well as deletions, additions, and substitutions.
<!-- a2m in the results, I put analyses based on all as well as just on first attempt. If we do just the latter, then come back to change this para -->
